---
title: "Categorical Analysis Techniques"
author: "Brian Bruxvoort"
date: "2023-12-07"
output: html_document
---

# Load necssary packages, libraries, and data sets
## Libraries
```{r}
library(dplyr)
library(car)
library(tidyverse)
library(leaps)
library(PropCIs) # For prop test
library(mcprofile)
library(nnet)
library(MASS)
library(vcd)
#library(glmulti)
library(randomForest)
library(rpart)
library(rpart.plot)
library(Metrics)
library(readxl)
library(caret)
```

## Data
```{r}
placekick <- read.csv("Placekick.csv") # Categorical Analysis
tomato <- read.csv("TomatoVirus.csv")
diet <- read.csv("Fiber.csv")
wheat <- read.csv("wheat.csv")
stoplight <- read.csv("Stoplight.csv")
crab <- read.table("Horseshoe.txt", header = TRUE)
```

# Categorical Analysis **Note* This code should only be used for reference as there are flaws*
## Chapter 1
### Two binary variables and Confidence intervals for the difference of two probabilities
```{r}
c.table <- array( data = c (57, 142, 200688, 201087) , dim = c(2, 2),
dimnames = list("Treatment" = c("Vaccine", "Placebo"),
"Polio" = c( "Polio", "Polie Free")))
c.table
```

```{r}
c.table [1 ,1]
c.table [1 ,]
sum(c.table[1 ,])
rowSums(c.table )
pi.hat.table <- c.table / rowSums(c.table)
pi.hat.table
```

```{r}
alpha <- 0.05
pi.hat1 <- pi.hat.table [1 ,1]
pi.hat2 <- pi.hat.table [2 ,1]
```

```{r}
#Wald
var.wald <- pi.hat1 * (1 - pi.hat1)/sum(c.table[1 ,]) +
pi.hat2 * (1 - pi.hat2) / sum(c.table [2 ,])
pi.hat1 - pi.hat2 + qnorm(p = c(alpha / 2, 1- alpha / 2))*
sqrt(var.wald)
```

```{r}
#Agresti-Caffo
pi.tilde1 <- (c.table [1 ,1] + 1)/(sum(c.table [1 ,]) + 2)
pi.tilde2 <- (c.table [2 ,1] + 1)/(sum(c.table [2 ,]) + 2)
var.AC <- pi.tilde1 * (1 - pi.tilde1)/(sum(c.table [1 ,]) + 2)
pi.tilde2 * (1 -pi.tilde2)/(sum(c.table [2 ,]) + 2)
pi.tilde1 - pi.tilde2 + qnorm(p = c(alpha / 2, 1- alpha /2))*sqrt(var.AC)
```

```{r}
w1 = 251
n1 = 285
w2 = 48
n2 = 53
pihat1 = w1/n1
pihat2 = w2/n2
pibar = (w1-w2)/(n1-n2)
z = (pihat1-pihat2)/sqrt((pibar*(1-pibar)*((1/n1)+(1/n2))))
z
qnorm(0.975)

chisq = (w1)
```

### Test for the difference of two probabilities
```{r}
prop.test(x = c.table, conf.level = 0.95, correct = FALSE)
```

```{r}
chisq.test(x = c.table, correct = FALSE)
qchisq(.95, 1)
```

### Relative Risk
```{r}
c.table2 <- array (data = c (251, 48, 34, 5), dim = c (2, 2), dimnames = list(First = c( "Made","Missed"), Second = c( "Made" , "Missed")))
c.table2
```

```{r}
pi.hat.table <- c.table2 / rowSums (c.table2)
pi.hat.table

pi.hat1 <- pi.hat.table [1 ,1]
pi.hat2 <- pi.hat.table [2 ,1]

round(pi.hat1 / pi.hat2, 4)

round(1/(pi.hat1/pi.hat2), 4)
```

### Relative Risk
```{r}
alpha<- 0.05
n1 <- sum(c.table2[1,])
n2 <- sum(c.table2[2,])
var.log.rr <- (1-pi.hat1) / (n1*pi.hat1) + (1-pi.hat2) / (n2*pi.hat2)
ci <- exp(log(pi.hat1/pi.hat2) + qnorm(p = c(alpha/2, 1-alpha/2)) * sqrt(var.log.rr))

round(ci, 4)

rev(round(1/ci, 4))

var.log.rr
```

```{r}
cat("The sample relative risk is", round(pi.hat1/pi.hat2, 4), "\n \n")

alpha<- 0.05
n1 <- sum(c.table2[1,])
n2 <- sum(c.table2[2,])
ci <- exp(log(pi.hat1/pi.hat2) + qnorm(p = c(alpha/2, 1-alpha/2)) * sqrt((1-pi.hat1)/(n1*pi.hat1) + (1-pi.hat2)/(n2*pi.hat2)))

round(ci, 4)

rev(round(1/ci, 4))
```

```{r}
(1-pi.hat1)/(1-pi.hat2)

exp(log((1-pi.hat1)/(1-pi.hat2)) + qnorm(p = c(alpha/2, 1-alpha/2)) * sqrt((pi.hat1)/(n1*(1-pi.hat1)) +(pi.hat2)/(n2*(1-pi.hat2))))
```

### Odds Ratio
```{r}
c.table3 <- array( data = c (57, 142, 200688, 201087) , dim = c(2, 2),
dimnames = list("Treatment" = c("Vaccine", "Placebo"),
"Polio" = c( "Polio", "Polie Free")))
c.table3
```

```{r}
OR.hat <- c.table3[1,1] * c.table3[2,2]/(c.table3[2 ,1] * c.table3[1 ,2])
round(OR.hat, 4)
round(1/OR.hat, 4)
```

```{r}
alpha <- 0.05
var.log.or <- 1/c.table3[1,1] + 1/c.table3[1,2] + 1/c.table3[2,1] + 1 /c.table3[2,2]
OR.CI <- exp(log(OR.hat) + qnorm(p = c(alpha / 2, 1-alpha / 2)) * sqrt(var.log.or))
round(OR.CI, 4)
rev(round(1/OR.CI, 4))
```

## Chapter 2
### 2.2.1 Parameter Estimation
```{r}
mod.fit <- glm(formula = good ~ distance, family = binomial(link = logit), data = placekick)
mod.fit
```

```{r}
mod.fit$coefficients
```

```{r}
summary(object = mod.fit)
```

```{r}
linear.pred <- mod.fit$coefficients[1] + mod.fit$coefficients[2]*20
exp(linear.pred)/(1+exp(linear.pred))
as.numeric(exp(linear.pred)/(1+exp(linear.pred)))
```

```{r}
linear.pred <- mod.fit$coefficients[1] + mod.fit$coefficients[2]*50
exp(linear.pred)/(1+exp(linear.pred))
as.numeric(exp(linear.pred)/(1+exp(linear.pred)))
```

```{r}
curve(expr = exp(mod.fit$coefficients[1] + mod.fit$coefficients[2]*x) / (1 + exp(mod.fit$coefficients[1] + mod.fit$coefficients[2]*x)), col = "red", xlim = c(18, 66), ylab = expression(hat(pi)), xlab = "Distance", main = "estimated probability of success for a placekick")
```

```{r}
mod.fit2 <- glm(formula = good ~ change + distance, family = binomial(link = logit), data = placekick)
summary(mod.fit2)
```

```{r}
vcov(mod.fit)

vcov(mod.fit)[2,2]
```

```{r}
summary(object = mod.fit)
```

```{r}
summary(mod.fit)$coefficients[,2]^2
```

```{r}
w<-aggregate(good ~ distance, data = placekick, FUN = sum)
n<-aggregate(good ~ distance, data = placekick, FUN = length)
w.n <- data.frame(distance = w$distance, success = w$good, trails = n$good, proportion = round(w$good/n$good, 4))
head(w.n)
```

```{r}
mod.fit.bin <- glm(formula = success/trails ~ distance, weight = trails, family = binomial(link = logit), data = w.n)
```

### 2.2.2 Hypothesis tests for regression parameters
```{r}
mod.fit2 <- glm(formula = good ~ change + distance, family = binomial(link = logit), data = placekick)
summary(mod.fit2)
```

```{r, message=FALSE, warning=FALSE}
Anova(mod = mod.fit2, test = "LR")
```

```{r}
anova(object = mod.fit2, test = "Chisq")
```

```{r}
mod.fit.Ho <- glm(formula = good ~ distance, family = binomial(link = logit), data = placekick)
anova(mod.fit.Ho, mod.fit2, test = "Chisq")
```

```{r}
mod.fit.Ho <- glm(formula = good ~ 1, family = binomial(link = logit), data = placekick)
mod.fit.Ha <- glm(formula = good ~ change, family = binomial(link = logit), data = placekick)
anova(mod.fit.Ho, mod.fit.Ha, test = "Chisq")
```

```{r}
pi.hat.Ho <- mod.fit.Ho$fitted.values
pi.hat.Ha <- mod.fit.Ha$fitted.values
y <- placekick$good
stat <-- 2*sum(y*log(pi.hat.Ho/pi.hat.Ha) + (1-y)*log((1-pi.hat.Ho)/(1-pi.hat.Ha)))
stat
pvalue<- 1-pchisq(q= stat, df=1)
data.frame(stat, pvalue)
```

### 2.2.3 Odds Ratio


### 2.2.4 Probability of Success
```{r}
predict.data <- data.frame(distance = c(20 ,30), change =c(1, 1))
predict.data
```

```{r}
alpha <- 0.05
linear.pred <- predict(object = mod.fit2, newdata = predict.data, type = "link", se = TRUE)
CI.lin.pred.x20 <- linear.pred$fit[1] + qnorm(p =c(alpha / 2, 1 - alpha / 2)) * linear.pred$se[1]
CI.lin.pred.x30 <- linear.pred$fit[2] + qnorm(p =c(alpha / 2, 1 - alpha / 2)) * linear.pred$se[2]
round(exp(CI.lin.pred.x20) / (1 + exp(CI.lin.pred.x20)), 4)
# CI for distance = 20
round(exp(CI.lin.pred.x30) / (1 + exp(CI.lin.pred.x30)), 4)
```

```{r}
K <- matrix(data = c(1, 20), nrow = 1, ncol = 2)
K
```

```{r}
linear.combo <- mcprofile(object = mod.fit, CM = K)
ci.logit.profile <- confint(object = linear.combo, level = 0.95)
ci.logit.profile
```

```{r}
names(ci.logit.profile)
exp(ci.logit.profile$confint)/(1 + exp(ci.logit.profile$confint))
```

```{r}
w <- aggregate(good ~ distance, data = placekick, FUN = sum)
n <- aggregate(good ~ distance, data = placekick, FUN = length)
w.n <- data.frame(distance = w$distance, success = w$good, trails = n$good, proportion = round(w$good/n$good, 4))
head(w.n)
```

```{r}
plot(x = w$distance, y = w$good/n$good, xlab="Distance (yards)", ylab = "Estimated probability", panel.first = grid(col = "gray", lty = "dotted"))
curve(expr = predict(object = mod.fit, newdata = data.frame(distance = x), type = "response"), col = "red", add = TRUE, xlim = c(18, 66))
```

```{r}
curve(expr = exp(mod.fit$coefficients[1] + mod.fit$coefficients[2] * x) / (1 + exp(mod.fit$coefficients[1] + mod.fit$coefficients[2] * x)), col = "red" , xlim = c(18, 66), ylab = expression(hat(pi)), xlab = "Distance", main = "Estimated probability of success for a placekick", panel.first = grid())
```

```{r}
symbols(x = w$distance, y = w$good / n$good, circles = sqrt(n$good), inches = 0.5, xlab = "Distance(yards)", ylab = "Estimated probability", panel.first = grid(col ="gray", lty = "dotted"))
curve(expr = predict(object = mod.fit, newdata = data.frame(distance = x), type = "response"), col = "red", add = TRUE, xlim = c(18, 66))
```

```{r}
ci.pi <- function(newdata, mod.fit.obj, alpha) {
  linear.pred<- predict(object = mod.fit.obj, newdata = newdata, type = "link", se = TRUE)
  CI.lin.pred.lower <- linear.pred$fit - qnorm(p = 1 - alpha/2) * linear.pred$se
  CI.lin.pred.upper <- linear.pred$fit + qnorm(p = 1 - alpha/2) * linear.pred$se
  CI.pi.lower <- exp(CI.lin.pred.lower) / (1 + exp(CI.lin.pred.lower))
  CI.pi.upper <- exp(CI.lin.pred.upper) / (1 + exp(CI.lin.pred.upper))
  list(lower = CI.pi.lower, upper = CI.pi.upper)
}

ci.pi(newdata = data.frame(distance = 20), mod.fit.obj = mod.fit, alpha = 0.05)
```

```{r}
curve(expr = ci.pi(newdata = data.frame(distance = x), mod.fit.obj = mod.fit, alpha = 0.05)$lower, col = "blue", lty = "dotdash", add = TRUE, xlim = c(18, 66))
curve(expr = ci.pi(newdata = data.frame(distance = x), mod.fit.obj = mod.fit, alpha = 0.05)$upper, col = "blue", lty = "dotdash", add = TRUE, xlim = c(18, 66))
legend(x = 20, y = 0.4, legend = c("Logistic regression model", "95% individual C.I."), lty = c("solid", "dotdash"), col = c("red", "blue"), bty = "n")
```


### 2.2.5 Interaction Terms
```{r}
mod.fit.Ha <- glm(good ~ distance + wind + distance:wind, family = binomial(link = logit), data = placekick)
summary(mod.fit.Ha)
```

```{r}
mod.fit.Ho <- glm(good ~ distance + wind, family = binomial(link = logit), data = placekick)
anova(mod.fit.Ho, mod.fit.Ha, test = "Chisq")
```

```{r}
curve(expr = predict(object = mod.fit.Ho, newdata = data.frame(distance = x, wind = 0), type = "response"), col = "red", lty = "solid", xlim = c(20,60), ylim = c(0,1), ylab = "Estimated probability", main = "Without Interaction", xlab = "Distance", panel.first = grid(col = "gray", lty = "dotted"), cex.main = 0.9, lwd = 1)

curve(expr = predict(object = mod.fit.Ho, newdata = data.frame(distance = x, wind = 1), type = "response"), col = "blue", lty = "dotdash", lwd = 1, add = TRUE)

legend

curve(expr = predict(object = mod.fit.Ha, newdata = data.frame(distance = x, wind = 0), type = "response"), col = "red", lty = "solid", xlim = c(20,60), ylim = c(0,1), ylab = "Estimated probability", main = "Without Interaction", xlab = "Distance", panel.first = grid(col = "gray", lty = "dotted"), cex.main = 0.9, lwd = 1)

curve(expr = predict(object = mod.fit.Ha, newdata = data.frame(distance = x, wind = 1), type = "response"), col = "blue", lty = "dotdash", lwd = 1, add = TRUE)
```

```{r}
curve(expr = exp(mod.fit.Ha$coefficients[1] + mod.fit.Ha$coefficients[2] * x) / (1 + exp(mod.fit.Ha$coefficients[1] + mod.fit.Ha$coefficients[2] * x)), col = "red" , xlim = c(18, 66), ylab = expression(hat(pi)), xlab = "Distance", main = "Estimated probability of success for a placekick", panel.first = grid())
```

```{r}
Anova(mod.fit.Ha, test = "LR")
```

```{r}
beta.hat <- mod.fit.Ha$coefficients[2:4]
c <- 1
distance <- seq(from = 20, to = 60, by = 10)
OR.wind <- exp(c*(beta.hat[2] + beta.hat[3] * distance))
cov.mat <- vcov(mod.fit.Ha)[2:4,2:4]
# Var(beta^_2 + distance * beta^_3)
var.log.OR <- cov.mat[2,2] + distance^2 * cov.mat[3,3] + 2 * distance * cov.mat[2, 3]
ci.log.OR.low <- c * (beta.hat[2] + beta.hat[3] * distance) - c * qnorm(p = 0.975) * sqrt(var.log.OR)
ci.log.OR.up <- c * (beta.hat[2] + beta.hat[3] * distance) + c * qnorm(p = 0.975) * sqrt(var.log.OR)
round(data.frame(distance = distance, OR.hat = 1 / OR.wind, OR.low = 1 / exp(ci.log.OR.up), OR.up = 1 / exp(ci.log.OR.low)),2)
```

```{r}
c <- 10
wind <- 0:1
OR.dist <- exp(c*(beta.hat[1] + beta.hat[3]*wind))
#Estimated OR
var.log.OR <- cov.mat [1,1] + wind^2*cov.mat[3,3] + 2*wind*cov.mat[1,3]
ci.log.OR.low <- c*(beta.hat[1] + beta.hat[3]*wind) - c*qnorm(p = 0.975) * sqrt(var.log.OR)
ci.log.OR.up <- c*(beta.hat[1] + beta.hat[3]*wind) + c*qnorm(p = 0.975) * sqrt(var.log.OR)
data.frame(OR.dist, OR.low = exp(ci.log.OR.low), OR.up = exp(ci.log.OR.up))
```


```{r}
beta.hat <- mod.fit.Ha$coefficients[2:4]
c<-1
distance <- seq(from = 20, to = 60, by = 10)
OR.wind <- exp(c*(beta.hat[2] + beta.hat[3] * distance))
cov.mat <- vcov(mod.fit.Ha)[2:4,2:4]
#Var ( beta ^ _ 2 + distance * beta ^ _ 3)
var.log.OR <- cov.mat[2,2] + distance^2 * cov.mat[3,3] + 2 * distance * cov.mat[2,3]
ci.log.OR.low <- c * (beta.hat[2] + beta.hat[3] * distance) - c * qnorm(p = 0.975) * sqrt(var.log.OR)
ci.log.OR.up <- c * (beta.hat[2] + beta.hat[3] * distance) + c * qnorm(p = 0.975) * sqrt(var.log.OR)
round(data.frame(distance = distance, OR.hat = 1 / OR.wind, OR.low = 1 / exp(ci.log.OR.up), OR.up = 1 / exp(ci.log.OR.low)), 2)
```

```{r}
c <- 10 # 10 - yard increment
wind <- 0:1 # Examine wind for 0 and 1
OR.dist <- exp(c*(beta.hat[1] + beta.hat[3] * wind))
#Estimated OR
var.log.OR <- cov.mat[1,1] + wind ^2 * cov.mat[3,3] + 2 * wind * cov.mat[1,3]
# Var ( beta ^_ 2 + distance * beta ^ _ 3)
ci.log.OR.low <- c * (beta.hat[1] + beta.hat[3] * wind) - c * qnorm(p = 0.975) * sqrt(var.log.OR)
ci.log.OR.up <- c * (beta.hat[1] + beta.hat[3] * wind) + c * qnorm(p = 0.975) * sqrt(var.log.OR)
data.frame(OR.dist, OR.low = exp(ci.log.OR.low),
OR.up = exp(ci.log.OR.up))
```

### 2.2.6
#### Recoding a variable with reference to one variable
```{r}
set1 <- data.frame(cat.var = c("D", "A", "A", "B", "D", "C"))
set1

class(set1$cat.var)
levels(set1$cat.var)
#contrasts(set1$cat.var)
```

```{r}
#cat.var2 <- relevel(x = set1$cat.var, ref = "D")
#set2 <- data.frame(set1, cat.var2)
#set2
```

```{r}
#set1$cat.var2 <- relevel(x = set1$cat.var, ref = "D")
#set1

#levels(set1$cat.var2)
#contrasts(set1$car.var2)
```

```{r}
class(tomato$Control)
levels(tomato$Control)
#contrasts(tomato$Control)
```

```{r}
levels(tomato$Infest)
class(factor(tomato$Infest))
levels(factor(tomato$Infest))
#contrasts(factor(tomato$Infest))
```

```{r}
tomato$Infest <- factor(tomato$Infest)
class(tomato$Infest)
```

```{r}
mod.fit <- glm(formula = Virus8/Plants ~ Infest + Control, family = binomial(link = logit), data = tomato, weight = Plants)
summary(mod.fit)
```

```{r}
mod.fit.inter <- glm(formula = Virus8/Plants ~ Control + Infest:Control, family = binomial(link = logit), data = tomato, weight = Plants)
summary(mod.fit.inter)
```

```{r}
mod.fit <- glm(formula = Virus8/Plants ~ Infest + Control, family = binomial(link = logit), data = tomato, weight = Plants)
summary(mod.fit)
```

```{r}
Anova(mod.fit.inter)
```

```{r}
mod.fit$xlevels
```

```{r}
exp(mod.fit$coefficients[3:4])
exp(mod.fit$coefficients[4] - mod.fit$coefficients[3])
```

```{r}
K <- matrix(data = c(0, 0, 1, 0, 0, 0, 0, 1), nrow = 2, ncol = 4, byrow = TRUE)
linear.combo <- mcprofile(object = mod.fit, CM = K)
ci.log.OR <- confint(object = linear.combo, level = 0.95, adjust = "none")
ci.log.OR
```

```{r}
comparison <- c("C vs B", "N vs B")
data.frame(comparison, OR = exp(ci.log.OR$confint))
```

```{r}
save.wald <- wald(object = linear.combo)
ci.logit.wald <- confint(object = save.wald, level = 0.95, adjust = "none")
data.frame(comparison, lower = exp(ci.logit.wald$confint[,1]), upper = exp(ci.logit.wald$confint[,2]))
```

```{r}
#Profile likelihood ratio interval
exp(confint(object = mod.fit, parm = c("ControlC", "ControlN"), level = 0.95))
```

```{r}
# Wald interval
exp(confint.default(object = mod.fit, parm = c("ControlC", "ControlN"), level = 0.95))
```

```{r}
K <- matrix(data = c(0, 0, -1, 1), nrow = 1, ncol = 4, byrow = TRUE)
linear.combo <- mcprofile(object = mod.fit, CM = K)
ci.log.OR <- confint(object = linear.combo, level = 0.95, adjust = "none")
ci.log.OR
```

```{r}
data.frame(comparison = "N vs C", OR = exp(ci.log.OR$confint))
save.wald <- wald(object = linear.combo)
ci.logit.wald <- confint(object = save.wald, level = 0.95, adjust = "none")
data.frame(lower = exp(ci.logit.wald$confint[,1]), upper = exp(ci.logit.wald$confint[,2]))
```

```{r}
beta.hat <- mod.fit$coefficients[-1]
exp(beta.hat[3] - beta.hat[2])
cov.mat <- vcov(mod.fit)[2:4, 2:4]
var.N.C <- cov.mat[3,3] + cov.mat[2,2] - 2*cov.mat[3,2]
CI.betas <- beta.hat[3] - beta.hat[2] + qnorm(p = c(0.025, 0.975)) * sqrt(var.N.C)
exp(CI.betas)
```

```{r}
mod.fit2 <- glm(formula = Virus8/Plants ~ Infest + Control, family = binomial(link = logit), data = tomato, weight = Plants)
summary(mod.fit2)
```

```{r}
# Wald interval
exp(confint.default(object = mod.fit2, parm = c("Control.reorderB", "Control.reorderN"), level = 0/95))

c(1/2.8629521, 1/1.707073)
```

```{r}
# Profile likelihood ratio interval
#exp(confint(object = mod.fit2, parm = c("Control.reorderB", "Control.reorderN"), level = 0.95))
```

```{r}
K <- matrix(data = c(0, 0, 0, 1, 0, 0, 0, 0, 0, 1, 0, 1, 0, 0, 1, 0, 0, 0, 0, 0, 1, 0, 1, 0, 0, 0, -1, 1, 0, 0, 0, 0, -1, 1, -1, 1), nrow = 6, ncol = 6, byrow = TRUE)
#linear.combo <- mcprofile(object = mod.fit.inter, Cm = K)
#ci.log.OR <- confint(object = linear.combo, level - 0.95, adjust = "none")
#ci.log.OR
```

```{r}
comparison <- c("N vs B", "N vs B", "C vs B", "C vs B", "N vs C", "N vs C")
data.frame(Infest2 = c(0, 1, 0, 1, 0, 1), comparison, OR = round(exp(ci.log.OR$estimate), 2), OR.CI = round(exp(ci.log.OR$confint), 2))
```

```{r}
save.wald <- wald(object = linear.combo)
ci.logit.wald <- confint(object = save.wald, level = 0.95, adjust = "none")
data.frame(Infest2 = c(0, 1, 0, 1, 0, 1), comparison, OR = round(exp(ci.log.OR$estimate), 2), lower = round(exp(ci.logit.wald$confint[,1]), 2), upper = round(exp(ci.logit.wald$confint[,2]), 2))
```

## Chapter 3
### 3.1 Multinomial Probability Distribution
```{r}
pi.j <- c(0.25, 0.35, 0.2, 0.1, 0.1)
set.seed(2195)
n.j <- rmultinom(n = 1, size = 1000, prob = pi.j)
data.frame(n.j, pihat.j = n.j/1000, pi.j)
```

```{r}
n.j <- rmultinom(n = 6, size = 1000, prob = pi.j)
n.j

n.j/1000
```

### 3.2 Contingency Tables and Inference Procedures
```{r}
pi.ij <- c(0.2, 0.3, 0.2, 0.1, 0.1, 0.1)
pi.table <- array(data = pi.ij, dim = c(2,3), dimnames = list(X = 1:2, Y = 1:3))
pi.table
```

```{r}
set.seed(9812)
save <- rmultinom(n = 1, size = 1000, prob = pi.ij)
c.table_1 <- array(data = save, dim = c(2,3), dimnames = list(X = 1:2, Y = 1:3))
c.table_1
c.table_1/sum(c.table_1)
```

```{r}
pi.cond <- pi.table/rowSums(pi.table)
pi.cond
```

```{r}
set.seed(8111)
save1 <- rmultinom(n = 1, size = 400, prob = pi.cond[1,])
save2 <- rmultinom(n = 1, size = 600, prob = pi.cond[2,])
c.table_2 <- array(data = c(save1[1], save2[1], save1[2], save2[2], save1[3], save2[3]), dim = c(2,3), dimnames = list(X = 1:2, Y = 1:3))
c.table_2

rowSums(c.table_2)

c.table_2/rowSums(c.table_2)
round(c.table_1/rowSums(c.table_1), 4)
```

```{r}
diet$fiber <- factor(x = diet$fiber, levels = c("none", "bran", "gum", "both"))
diet$bloat <- factor(x = diet$bloat, levels = c("none", "low", "medium", "high"))
diet.table <- xtabs(formula = count ~ fiber + bloat, data = diet)
diet.table
```

```{r}
ind.test <- chisq.test (x = diet.table, correct = FALSE)
ind.test
```

```{r}
class(diet.table)
summary(diet.table)
qchisq(p = 0.95, df = 9)
```

```{r}
ind.test$expected
```

### 3.3 Nominal Response Models
```{r}
levels(wheat$type)

mod.fit.nom.density <- multinom(formula = type ~ class + density + hardness + size + weight + moisture, data = wheat)
summary(mod.fit.nom.density)
```

```{r}
#lwd.po <- 4

#curve(expr = plogis(q = mod.fit.nom.density$zeta[1] - mod.fit.nom.density$coefficients*x), col = "green", type = "l", xlim =c(min(wheat$density[wheat$type.order == "Scab"]), max(wheat$density[wheat$type.order == "Scab"])), add = TRUE, lty = "dotdash", lwd = lwd.po, n = 1000)

#curve(expr = plogis(q = mod.fit.nom.density$zeta[2] - mod.fit.nom.density$coefficients*x) - plogis(q = mod.fit.nom.density$zeta[1] - mod.fit.nom.density$coefficients*x), col = "red", type = "l", xlim =c(min(wheat$density[wheat$type.order == "Sprout"]), max(wheat$density[wheat$type.order == "Sprout"])), add = TRUE, lty = "longdash", lwd = lwd.po, n = 1000)

#curve(expr = 1 - plogis(q = mod.fit.nom.density$zeta[2] - mod.fit.nom.density$coefficients*x), col = "black", type = "l", xlim =c(min(wheat$density[wheat$type.order == "Healthy"]), max(wheat$density[wheat$type.order == "Healthy"])), add = TRUE, lty = "solid", lwd = lwd.po, n = 1000)

#legend(x = 1.4, y = 0.8, legend = c("Healthy", "Sprout", "Scab"), lty = c("solid", "longdash", "dotdash"), col = c("black", "red", "green"), bty = "n", lwd = c(2,2,2), seg.len = 4)
```

```{r}
density.values <- seq(from = 0.8, to = 1.6, by = 0.1)
data.frame(density.values, round(predict(object = mod.fit.nom.density, newdata = data.frame(density = density.values), type = "probs"), 2))
```

```{r}
wheat2 <- data.frame(kernal = 1:nrow(wheat), wheat[,2:6], class.new = ifelse(test = wheat$class == "hrw", yes = 0, no = 1))
head(wheat2)

wheat.colors <- ifelse(test = wheat$type == "healthy", yes = "black", no = ifelse(test = wheat$class == "Sprout", yes = "red", no = "green"))

wheat.lty
```

### 3.3.1 Odds Ratios
```{r}
sd.wheat <- apply(X = wheat[,-c(1,7)], MARGIN = 2, FUN = sd)
c.value <- c(1, sd.wheat)
round(c.value, 2)
beta.hat2 <- coefficients(mod.fit)[1, 2:7]
beta.hat3 <- coefficients(mod.fit)[2, 2:7]
round(exp(c.value*beta.hat2), 2)
round(1/exp(c.value*beta.hat2), 2)
round(exp(c.value*beta.hat3), 2)
round(1/exp(c.value*beta.hat3), 2)
```

```{r}
conf.beta <- confint(object = mod.fit, level = 0.95)
conf.beta
```

```{r}
ci.OR2 <- exp(c.value*conf.beta[2:7, 1:2, 1])
ci.OR3 <- exp(c.value*conf.beta[2:7, 1:2, 2])
round(data.frame(low = ci.OR2[,1], up = ci.OR2[,2]), 2)
round(data.frame(low = 1/ci.OR2[,2], up = 1/ci.OR2[,1]), 2)
round(data.frame(low = ci.OR3[,1], up = ci.OR3[,2]), 2)
round(data.frame(low = 1/ci.OR3[,2], up = 1/ci.OR3[,1]), 2)
```

### 3.3.2 Contingency Tables
```{r}
mod.fit.nom <- multinom(formula = bloat ~ fiber, weights = count, data = diet)
summary(mod.fit.nom)
```

```{r}
Anova(mod.fit.nom)
```

### 3.4 Ordinal Response Models
```{r}
levels(wheat$type)
wheat$type.order <- factor(wheat$type, levels = c("scab", "Sprout", "Healthy"))
levels(wheat$type.order)
```

```{r}
mod.fit.ord <- polr(type.order ~ class + density + hardness + size + weight + moisture, data = wheat, method = "logistic")
summary(mod.fit.ord)
```

```{r}
Anova(mod.fit.ord)
```

```{r}
pi.hat.ord <- predict(object = mod.fit.ord, type = "probs")
head(pi.hat.ord)
head(predict(object = mod.fit.ord, type = "class"))
```

### 3.4.1 Odds Ratios
```{r}
sd.wheat <- apply(X = wehat[,-c(1,7)], MARGIN = 2, FUN = sd)
round(c.value, 2)
c.value <- c(1, sd.wheat)
round(exp(c.value * (-mod.fit.ord$coefficients)), 2)
round(1/exp(c.value * (-mod.fit.ord$coefficients)), 2)
```

```{r}
conf.beta <- confint(object = mod.fit.ord, level = 0.95)
ci <- exp(c.value*(-conf.beta))
round(data.frame(low = ci[,2], up = ci[,1]), 2)
round(data.frame(low = 1/ci[,1], up = 1/ci[,2]), 2)
```

### 3.4.2 Contingency Tables
```{r}
levels(diet$bloat)
mod.fot.ord <- polr(formula = bloat ~ fiber, weights = count, data = diet, method = "logistic")
summary(mod.fit.ord)
```

```{r}
Anova(mod.fit.ord)
```

### 3.4.3 Non-proportional odds model
```{r}
vlgm()
```

### 3.4
```{r}
beta <- c(0, 2, 4, 2)
x.range <- c(-5, 3)
par(mfrow = c(1,1))
setEPS()
postscript("3-4.eps")
curve(expr = plogis(q = beta[1] + beta[4] * x), xlim = x.range, ylab = expression(P(Y<=j)), xlab = expression(x[1]), main = "Cumulative probabilites for Y", lwd =2)
curve(expr = plogis(q = beta[2] + beta[4]*x), add = TRUE, lty = "dashed", col ="red", lwd = 2)
curve(expr = plogis(q = beta[3] + beta[4]*x), add = TRUE, lty = "dotted", col ="blue", lwd = 2)
legend(x = -0.5, y = 0.9, legend = c(expression(P(Y<=1)), expression(P(Y<=2)), expression(P(Y<=3))), lty = c("solid", "dashed", "dotted", "dotdash"), col = c("black", "red", "blue"), bty="n", lwd =2)
```

## Chapter 4
### 4.1 Poisson Model for Count Data
```{r}
mean(stoplight$vehicles)
var(stoplight$vehicles)
table(stoplight$vehicles)
rel.freq <- table(stoplight$vehicles) / length(stoplight$vehicles)
rel.freq2 <- c(rel.freq, rep(0, times =7))
```

```{r}
y <- 0:15
prob <- round(dpois(x=y, lambda = mean(stoplight$vehicles)))
data.frame(y, prob, rel.freq = rel.freq2)
```

```{r}
plot(x = y-0.1, y = prob, type = "h", ylab = "Probability", xlab = "Number of Vechicles", lwd = 2, xaxt = "n")
axis(side = 1, at = 0:15)
lines(x = y+0.1, y = rel.freq2, type = "h", lwd = 2, lty = "solid", col = "red")
abline(h = 0)
legend(x = 9, y = 0.15, legend = c("Poisson", "Observed"), lty = c("solid", "solid"), lwd = c(2,2), col = c("black", "red"), bty = "n")
```

```{r}
alpha <- 0.05
n <- length(stoplight$vehicles)
mu.hat <- mean(stoplight$vehicles)
lower <- (mu.hat + (qnorm(p = c(alpha/2, 1-alpha/2)))^2) - (qnorm(p = c(alpha/2, 1-alpha/2))) * sqrt((mu.hat-((qnorm(p = c(alpha/2, 1-alpha/2)))^2)/(4*n))/n)
upper <- (mu.hat + (qnorm(p = c(alpha/2, 1-alpha/2)))^2) + (qnorm(p = c(alpha/2, 1-alpha/2))) * sqrt((mu.hat-((qnorm(p = c(alpha/2, 1-alpha/2)))^2)/(4*n))/n)
interval
```

#4.2.2 Parameter estimation and inference
```{r}
mod.fit <- glm(formula = satellite ~ width, data = crab, family = poisson(link = log))
summary(mod.fit)
```

```{r}
plot(x = crab$width, y = crab$satellite, xlab = "Width (cm)", ylab = "Number of Satellites", main = "Horseshoe crab data set with Poisson regression model fit", panel.first = grid())
curve(expr = exp(mod.fit$coefficients[1] + mod.fit$coefficients[2]*x), col = "red", add = TRUE, lty = "solid")
ci.mu <- function(newdata, mod.fit.obj, alpha) {
  lin.pred.hat <- predict(object = mod.fit.obj, newdata = newdata, type = "link", se = TRUE)
  lower <- exp(exp.pred.hat$fit - qnorm(1 - alpha/2) * lin.pred.hat$se)
  upper <- exp(exp.pred.hat$fit + qnorm(1 - alpha/2) * lin.pred.hat$se)
  list(lower = lower, upper = upper)
}
```

```{r}
ci.mu(newdata = data.frame(width = 23), mod.fit.obj = mod.fit, alpha = 0.95)
```

```{r}
curve(expr = ci.mu(newdata = data.frame(width = x), mod.fit.obj = mod.fit, alpha = 0.95)$lower, col = "blue", add = TRUE, lty = "dotdash")
curve(expr = ci.mu(newdata = data.frame(width = x), mod.fit.obj = mod.fit, alpha = 0.95)$upper, col = "blue", add = TRUE, lty = "dotdash")
legend(x = 21, y = 14, legend = c("Poisson regression model", "95% individual C.I."), bty = "n", lty = c("solid", "dotdash"), col = c("red", "blue"))
```

```{r}
groups <- ifelse(test = crab$width<23.25, yes = 1, no = ifelse(test = crab$width<24.25, yes = 2, no =ifelse(test = crab$width<25.25, yes = 3, no =ifelse(test = crab$width<26.25, yes = 4, no =ifelse(test = crab$width<27.25, yes = 5, no =ifelse(test = crab$width<28.25, yes = 6, no =ifelse(test = crab$width<29.25, yes = 7, no = 8)))))))
crab.group <- data.frame(crab, groups)
head(crab.group)
```

```{r}
ybar <- aggregate(formula = satellite ~ groups, data = crab, FUN = mean)
xbar <- aggregate(formula = width ~ groups, data = crab, FUN = mean)
data.frame(ybar, xbar$width)
```

```{r}
points(x = xbar$width, y = ybar$satellite, pch = 17, col = "darkgreen", cex = 2)
#first legend() call
```

```{r}
legend(x = 21, y = 14, legned = c("Poisson regression model", "95% individual C.I.", "sample mean"), bty = "n", lty = c("solid", "dotdash", NA), col = c("red", "blue", "darkgreen"), pch = c(NA, NA, 17))
```

```{r}
exp(mod.fit$coefficients[2])
100*(exp(mod.fit$coefficients[2]) - 1)
```

```{r}
c.unit <- sd(crab$width)
c.unit
100*(exp(c.unit*mod.fit$coefficients[2]) - 1)
```

```{r}
beta.ci <- confint(object = mod.fit, parm = "width", level = 0.95)
100*(exp(beta.ci) - 1)
```

```{r}
K <- matrix(data = c(0, 1), nrow = 1, ncol = 2)
linear.combo <- mcprofile(object = mod.fit, CM = K)
ci.beta <- confint(object = linear.combo, level = 0.95)
100*(exp(ci.beta$confint) - 1)
```

```{r}
beta.ci <- confint.default(object = mod.fit, parm = "width", level = 0.95)
beta.ci
exp(beta.ci)
100*(exp(beta.ci) - 1)
```

```{r}
vcov(mod.fit)
```

```{r}
beta.ci <- mod.fit$coefficients[2] + qnorm(p = c(0.025, 0.975)) * sqrt(vcov(mod.fit)[2,2])
100*(exp(beta.ci) - 1)
```

#4.2.4 Poisson Regression for Contingency Tables: LogLinear Models
```{r}
c.table4 <- array(data = c(251, 48, 34, 5), dim = c(2,2), dimnames = list(First = c("made", "missed"), Second = c("made", "missed")))
bird <- as.data.frame(as.table(c.table4))
bird
```

```{r}
mod.fit1 <- glm(formula = Freq ~ First + Second, data = bird, family = poisson(link = log))
summary(mod.fit1)
```

```{r}
predict(object = mod.fit1, newdata = bird, type = "response")
pearson.test <- chisq.test(x = c.table, correct = FALSE)
pearson.test$expected
```

```{r}
mod.fit2 <- glm(formula = Freq ~ First + Second + First:Second, data = bird, family = poisson(link = log))
summary(mod.fit2)
```

```{r}
predict(object = mod.fit2, newdata = bird, type = "response")
bird$Freq
```

```{r}
Anova(mod.fit2, test = "LR")
```

```{r}
anova(mod.fit1, mod.fit2, test = "Chisq")
```

```{r}
assocstats(c.table)
```

#4.3 Poisson Rate Regression
```{r}
total.sat <- aggregate(satellite ~ width, data = crab, FUN = sum)
numb.crab <- aggregate(satellite ~ width, data = crab, FUN = length)
rate.data <- data.frame(total.sat, numb.crab = numb.crab$satellite)
head(rate.data)
```

```{r}
mod.fit.rate <- glm(formula = satellite ~ width + offset(log(numb.crab)), data = rate.data, family = poisson(link = log))
summary(mod.fit.rate)
```

```{r}
predict(object = mod.fit.rate, newdata = data.frame(width = c(23, 23), numb.crab = c(1, 2)), type = "response")
```

```{r}
plot.char.numb <- as.numeric(names(table(rate.data$numb.crab)))
plot(x = rate.data$width, y = rate.data$satellite, xlab = "Width (cm)", ylab = "Number of satellites", type = "n", panel.first = grid(), main = "Horseshoe crab data set with Poisson regression model fir (rate data)")
for (t in plot.char.numb) {
  width.t <- rate.data$width[rate.data$numb.crab == plot.char.numb[t]]
  satellite.t <- rate.data$satellite[rate.data$numb.crab == plot.char.numb[t]]
  points(x = width.t, y = satellite.t, pch = as.character(plot.char.numb[t]), cex = 0.5, col = t)
  curve(expr = t * exp(mod.fit.rate$coefficients[1] + mod.fit.rate$coefficients[2]*x), xlim = c(min(width.t), max(width.t)), lty = "solid", col = t, add = TRUE)
}
```

```{r}
for (t in plot.char.numb) {
  width.t <- rate.data$width[rate.data$numb.crab == plot.char.numb[t]]
  satellite.t <- rate.data$satellite[rate.data$numb.crab == plot.char.numb[t]]
  points(x = width.t, y = satellite.t, pch = as.character(plot.char.numb[t]), cex = 0.5, col = t)
  curve(expr = t * exp(mod.fit.rate$coefficients[1] + mod.fit.rate$coefficients[2]*n), xlim = c(min(width.t), max(width.t)), lty = "solid", col = t, add = TRUE)
}
```

```{r}
Anova(mod.fit)
```

```{r}
Anova(mod.fit.rate)
```

## Chapter 5
### 5.1 Variable Selection
```{r}
search.1.aicc <- glmulti(y = good ~ ., data = placekick, fitfunction = "glm", level = 1, method = "h", crit = "aicc", family = binomial(link="logit"))
```

```{r}
aa <- weightable(search.1.aicc)
head(aa)
```

```{r}
cbind(model = aa[1:5,1], round(aa[1:5, c(2,3)], 3))
```

```{r}
set.seed(87112811)
search.gmarg.aicc <- glmulti(y = good ~ ., data = placekick, fitfunction = "glm", level = 2, marginality =TRUE, method = "g", crit = "aicc", family = binomial(link = "logit"))
```

```{r}
print(search.gmarg.aicc)
```

```{r}
set.seed(91051211)
head(weightable(search.gmarg.aicc))
```

```{r}
set.seed(11256012)
head(weightable(search.gmarg.aicc))
```

```{r}
empty.mod <- glm(good ~ 1, family = binomial(link = logit), data = placekick)
full.mod <- glm(good ~ ., family = binomial(link = logit), data = placekick)
```

```{r}
forw.sel <- step(object = empty.mod, scope = list(upper = full.mod), direction = "forward", k = log(nrow(placekick)), trace = TRUE)
```

```{r}
anova(forw.sel)
```

```{r}
forw.sel2 <- step(object = empty.mod, scope = list(upper = full.mod), direction = "forward", k = 2, trace = TRUE)
```

```{r}
anova(forw.sel2)
```

```{r}
search.1.bic <- glmulti(y = good ~ ., data = placekick, fitfunction = "glm", level = 1, method = "h", crit = "bic", family = binomial(link = "logit"))
print(search.1.bic)
```

```{r}
head(weightable(search.1.bic))
```

```{r}
plot(search.1.bic, type = "w")
```

```{r}
parms <- coef(search.1.bic)
colnames(parms) <- c("Estimate", "Variance", "n.Models", "Probability", "95%CI +/-")
round(parms, digits = 3)
```

```{r}
parm.ord <- parms[order(parms[,4], decreasing = TRUE),]
ci.parms <- cbind(lower = parm.ord[,1] - parms.ord[,5], upper = parm.ord[,1] - parms.ord[,5])
round(cbind(parms.ord[,1], ci.parms), digits = 3)
round(exp(cbind(OR = parms.ord[,1], ci.parms))[-1,], digits = 2)
```

```{r}
best.fit <- glm(good ~ distance + PAT, data = placekick, family = binomial(link = "logit"))
round(summary(best.fit)$coefficients, 2)
```

#5.2 Tools to Assesss Model Fit
```{r}
w <- aggregate(good ~ distance, data = placekick, FUN = sum)
n <- aggregate(good ~ distance, data = placekick, FUN = length)
w.n <- data.frame(distance = w$distance, success = w$good, trials = n$good, proportion = round(w$good/n$good, 4))
head(w.n)
```

```{r}
y <- 0.3
n <- 3
pi.hat <- 0.9
e <- round((y - n*pi.hat)/sqrt(n*pi.hat*(1-pi.hat)), 2)
prob.norm <- round(pnorm(q = e), 4)
prob.bin <- round(pbinom(q = y, size = n, prob = pi.hat), 4)
data.frame(y, e, prob.norm, prob.bin)
```

```{r}
mod.fit.bin <- glm(success/trails ~ distance, weights = trails, family = binomial(link = logit), data = w.n)
pi.hat <- predict(mod.fit.bin, type = "response")
p.res <- residuals(mod.fit.bin, type = "pearson")
s.res <- rstandard(mod.fit.bin, type = "pearson")
lin.pred <- mod.fit.bin$linear.predictors
w.n <- data.frame(w.n, pi.hat, p.res, s.res, lin.pred)
round(head(w.n), digits = 3)
```

```{r}
plot(x = w.n$distance, y = w.n$s.res, xlab = "Distance", ylab = "Standardized Pearson residuals", main = "Standardized residuals vs. X")
abline(h = c(3, 2, 0, -2, -3), lty = 3, col = "blue")
smooth.stand <- loess(s.res ~ distance, data = w.n, weights = trials)
order.dist <- order(w.n$distance)
lines(x = w.n$distance[order.dist], y = predict(smooth.stand)[order.dist], lty = 3, col = "red", lwd = 3)
```

```{r}
1 - pbinom(q = 775, size = 789, prob = 0.971)
```

# 5.3 Overdispersion
```{r}
groups <- ifelse(test = crab$width<23.25, yes = 1, no = ifelse(test = crab$width<24.25, yes = 2, no =ifelse(test = crab$width<25.25, yes = 3, no =ifelse(test = crab$width<26.25, yes = 4, no =ifelse(test = crab$width<27.25, yes = 5, no =ifelse(test = crab$width<28.25, yes = 6, no =ifelse(test = crab$width<29.25, yes = 7, no = 8)))))))
crab.group <- data.frame(crab, groups)
head(crab.group)
```

```{r}
ybar <- aggregate(formula = satellite ~ groups, data = crab, FUN = mean)
var.y <- aggregate(satellite ~ groups, data = crab, FUN = var)
group.name <- c("width <23.25", "23.25 <= width < 24.25", "24.25 <= width < 25.25", "25.25 <= width < 26.25", "26.25 <= width < 27.25", "27.25 <= width < 28.25", "28.25 <= width < 29.25", "29.25 <= width < 30.25")
data.frame(group.name, mean.sat = round(ybar$satellite, 2), var.sat = round(var.y$satellite, 2))
```

```{r}
mod.fit <- glm(satellite ~ width, data = crab, family = poisson(link = log))
summary(mod.fit)
```

```{r}
mod.fit$deviance / mod.fit$df.residual

round(c(1 + 2*sqrt(2/mod.fit$df.residual), 1 + 3*sqrt(2/mod.fit$df.residual)), 2)
```

```{r}
mu.hat <- mod.fit$fitted.values
stand.resid <- rstandard(model = mod.fit, type = "pearson")
plot(x = mu.hat, y = stand.resid, xlab = expression(hat(mu)), ylab = "Standard Pearson Residuals", ylim = c(min(c(-3, stand.resid)), max(c(3, stand.resid))))
abline(h = c(-3, -2, 0, 2, 3), lty = "dotted", col = "red")
```

```{r}
sum(abs(stand.resid) > 3)
length(stand.resid)
sum(abs(stand.resid) > 3) / length(stand.resid)
length(stand.resid)*2*(1 - pnorm(q = 3))
```

```{r}
mod.fit.quasi <- glm(satellite ~ width, data = crab, family = quasipoisson(link = log))
summary(mod.fit.quasi)
```

```{r}
sum.fit.quasi <- summary(mod.fit.quasi)
vcov(mod.fit.quasi)
vcov(mod.fit) * sum.fit.quasi$dispersion
```

```{r}
anova(mod.fit.quasi, test = "F")
Anova(mod.fit.quasi, test.statistic = "F")
```

```{r}
beta.ci <- confint.default(object = mod.fit.quasi, parm = "width", level = 0.95)
beta.ci
100*(exp(beta.ci) - 1)
```

```{r}
beta.ci <- confint(object = mod.fit.quasi, parm = "width", level = 0.95)
beta.ci
100*(exp(beta.ci)- 1)
```

```{r}
mod.fit.nb <- glm.nb(satellite ~ width, data = crab, link = log)
summary(mod.fit.nb)
class(mod.fit.nb)
```

```{r}
vcov(mod.fit.nb)
vcov(mod.fit.quasi)
vcov(mod.fit)
```

```{r}
sum.fit.nb <- summary(mod.fit.nb)
sum.fit <- summary(mod.fit)
std.err <- data.frame(Poisson = sum.fit$coefficients[,2], quasi = sum.fit.quasi$coefficients[,2], nb = sum.fit.nb$coefficients[,2])
std.err
```

```{r}
nb23 <- ci.mu(newdata = data.frame(width = 23), mod.fit.obj = mod.fit.nb, alpha = 0.05)
quasi23 <- ci.mu(newdata = data.frame(width = 23), mod.fit.obj = mod.fit.quasi, alpha = 0.05)
Pois23 <- ci.mu(newdata = data.frame(width = 23), mod.fit.obj = mod.fit, alpha = 0.05)
data.frame(type = c("Negative binomial", "Quasi-Poisson", "Poisson"), lower = round(c(nb23$lower, quasi23$lower, Pois23$lower), 2), upper = round(c(nb23$upper, quasi23$upper, Pois23$upper), 2))
```

```{r}
nb33 <- ci.mu(newdata = data.frame(width = 33), mod.fit.obj = mod.fit.nb, alpha = 0.05)
quasi33 <- ci.mu(newdata = data.frame(width = 33), mod.fit.obj = mod.fit.quasi, alpha = 0.05)
Pois33 <- ci.mu(newdata = data.frame(width = 33), mod.fit.obj = mod.fit, alpha = 0.05)
data.frame(type = c("Negative binomial", "Quasi-Poisson", "Poisson"), lower = round(c(nb33$lower, quasi33$lower, Pois33$lower), 2), upper = round(c(nb33$upper, quasi33$upper, Pois33$upper), 2))
```

```{r}
set1 <- data.frame(res.sq = residuals(object = mod.fit, type = "response")^2, mu.hat = mod.fit$fitted.values)
fit.lin <- lm(res.sq ~ mu.hat, data = set1)
fit.quad <- lm(res.sq ~ mu.hat + I(mu.hat^2), data = set1)
summary(fit.quad)
```

```{r}
plot(z = set1$mu.hat, y = set1$res.sq, xlab = expression(hat(mu)), ylab = "Squared Residual")
curve(expr = predict(object = fit.lin, newdata = data.frame(mu.hat = x), type = "response"), col = "blue", add = TRUE, lty = "solid")
curve(expr = predict(object = fit.quad, newdata = data.frame(mu.hat = x), type = "response"), col = "red", add = TRUE, lty = "solid")
legend(x = 50, y =1000, legend = c("Linear", "Quadratic"), col = c("red", "blue"), lty = c("solid", "dashed"), bty = "n")
```
